''''

---
title: "2019_31_Video_Game"
author: "Ryan Peiffer"
date: "7/29/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
options(scipen = 999)
```


This week's Tidy Tuesday is about video game data from Steam! 
First let's load the data into the workspace:


```{r load data, message=FALSE}
vg <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv")

head(vg)

```


I definitely want to do something with the "owners" column but the current format is rough to deal with.
I'm going to create a new column representing the min of each bucket. This is certainly not a great assumption
but at least it gives me numbers to work with!


```{r fix owners, warning=FALSE}
vg <- vg %>% mutate(owners2 = as.numeric(gsub(",", "", substr(owners, 0, regexpr('\\.{2}', owners)-2))))

```


What games have been played the most total hours?


```{r most played}

most_played <- vg %>%
    mutate(total_playtime = owners2 * average_playtime) %>%
    arrange(desc(total_playtime)) %>%
    slice(1:20)
    
ggplot(most_played, aes(reorder(game, total_playtime), total_playtime)) +
    geom_bar(aes(fill = average_playtime), stat = "identity") +
    coord_flip()

```


What games have made the most money (assuming no DLC purchases)?


```{r highest grossing}

most_money <- vg %>%
    mutate(sales = owners2 * price) %>%
    arrange(desc(sales)) %>%
    slice(1:20)
    
ggplot(most_money, aes(reorder(game, sales), sales)) +
    geom_bar(aes(fill = price), stat = "identity") +
    coord_flip()
```

Let's see how much price influences total downloads.
There are 200 unique prices in the data... I'm going to do some rounding to make it better to work with.

```{r rounded prices}

my_prices <- c(0, 0.49, 0.99, 4.99, 9.99, 14.99, 19.99, 49.99, 99.99, 1000)

round_prices <- vg %>%
    filter(!is.na(price)) %>%
    mutate(round_price = cut(price, breaks = my_prices))

price_buckets <- round_prices %>%
    count(round_price)

ggplot(price_buckets, aes(round_price, n)) +
    geom_bar(stat = "identity")

```

Cool. Now let's see total owners and average owners per game by price bucket.

```{r price influence}

ggplot(round_prices, aes(round_price, owners2)) +
    geom_bar(stat = "sum")

#PUBG is throwing off the y axis... let's yank it out for now.
round_prices2 <- filter(round_prices, game != "PLAYERUNKNOWN'S BATTLEGROUNDS")

ggplot(round_prices2, aes(round_price, owners2)) +
    geom_bar(stat = "sum")

avg_players_by_price <- round_prices2 %>%
    group_by(round_price) %>%
    summarise(avg_players = mean(owners2))

ggplot(avg_players_by_price, aes(round_price, avg_players)) + 
    geom_bar(stat = "identity")

```



''''
