---
title: "Do Viewers like Michael Scott?"
author: "Ryan Peiffer"
date: "3/31/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Tidytuesday's 2020 week 12 challenge is on The Office! Everyone's favorite show
featuring everyone's favorite character, Michael Scott. Or is he? I'd like to 
take a look at how IMDB ratings relate to Michael's (verbal) presence.

Let's get started!

```{r, message=FALSE, warning=FALSE}
#load packages and read in dataset
library(tidyverse)
library(gghighlight)
library(schrute)

raw_data <- schrute::theoffice

raw_data
```

The Schrute package (great name) pulls in every line by every character in 
every episode. Awesome! For my purposes, I'm going to want to know how many
words each character is speaking. So let's manipulate this data.

```{r}
lines_manip <- raw_data %>%
    select(season, episode, character, text) %>%
    tidytext::unnest_tokens(word, text) %>%
    mutate(ep_key = paste0("s", season, "e", episode)) %>%
    add_count(character, ep_key) %>%
    rename(char_words = n) %>%
    add_count(ep_key) %>%
    rename(ep_words = n)

lines_sum <- lines_manip %>%
    select(ep_key, character, char_words, ep_words) %>%
    distinct()
```

The "lines_manip" dataset gives me a 570k row dataset on every word said during
the show. Cool, but not that necessary for what I'm currently doing. "lines_sum"
aggregates this up to be one row per character per episode, counting their
total words in the episode, and the overall total words in the episode. Below you
can see an example for season 1 episode 1.

```{r, echo=FALSE}
lines_sum %>%
    filter(ep_key == "s1e1") %>%
    arrange(desc(char_words)) 
```


Now that we have a basic dataset to work off of, let's start digging into the
World's Greatest Boss (according to coffee mugs). I want to plot the IMDB rating
for each epsiode compared to Michael's share of the total word count.

```{r}
episodes <- raw_data %>%
    mutate(ep_key = paste0("s", season, "e", episode)) %>%
    select(ep_key, episode_name, imdb_rating) %>%
    unique()

michael_words <- lines_sum %>%
    filter(character == "Michael") 

michael_ratings <- episodes %>%
    left_join(michael_words, by = "ep_key") %>%
    mutate(michael_share = char_words / ep_words) %>%
    replace_na(list(michael_share = 0)) %>%
    arrange(michael_share)

ggplot(michael_ratings, aes(x = michael_share, y = imdb_rating)) +
    geom_point() +
    theme_bw()
```

Some things immediately stand out here. First question: what's up with the big
line of dots on the left? Well, Michael is not in every episode! Let's take a step
aside and see how the average rating changes based on Michael's presence.

```{r}
michael_presence <- michael_ratings %>%
    mutate(michael_exists = if_else(michael_share == 0, "No", "Yes")) %>%
    group_by(michael_exists) %>%
    summarise(avg_rating = mean(imdb_rating))

michael_presence %>%
    knitr::kable(digits = 1, col.names = c("Michael Exists", "Avg Rating"),
                 align = "c")
```

Looks like viewers do appreciate him being in the show! That's a good start.
Now that we have that taken care of, I want to dial in on the episodes that feature
Mike the Magic. Let's look at the same plot, but only showing episodes where Michael
speaks at least 1 word.

```{r, warning=FALSE}
michael_ratings_no0 <- episodes %>%
    left_join(michael_words, by = "ep_key") %>%
    mutate(michael_share = char_words / ep_words) %>%
    #replace_na(list(michael_share = 0))  >>> no longer replacing NAs!
    arrange(michael_share)

ggplot(michael_ratings_no0, aes(x = michael_share, y = imdb_rating)) +
    geom_point() +
    theme_bw()
```

Looks like there's a few outlier points. Let's see if we can figure out what's
up with these four:

```{r, warning=FALSE, echo=FALSE}
#manually building vector of the episodes I want to highlight
highlights <- c("s9e24", "s6e14", "s1e5", "s1e1")

michael_ratings_highlight <- michael_ratings_no0 %>%
    mutate(ep_label = paste0(ep_key, ": ", episode_name))

ggplot(michael_ratings_highlight, aes(x = michael_share, y = imdb_rating)) +
    geom_point() +
    gghighlight(ep_key %in% highlights, label_key = ep_label) +
    theme_bw()
    
```

Aha! This makes some sense. Hardcore fans of the show may have even guessed what
each of these data points are. Here's my brief insight into them:  

* **s1e1: Pilot:** The Pilot episode featured *a lot* of Mr. Scott and, like many pilot episodes, writers didn't quite find the magic just yet.  
* **s9e24: Finale:** Michael briefly returns for the finale, however viewers loved the finale for how well it celebrated *all* of their favorite characters.   
* **s6e14: The Banker:** This is a classic montage episode, which viewers tend to not like very much.  
* **s1e5: Basketball:** Not much to say here, other than Michael is heavily featured.

Now that we have our outliers understood, let's see if there's any correlation between Michael's share of the script and the IMDB rating. I'm going to remove the above four outlier episodes for this.

```{r}
michael_reg_data <- michael_ratings_highlight %>%
  filter(!ep_key %in% highlights) %>%
  filter(!is.na(michael_share))
  
ggplot(michael_reg_data, aes(x = michael_share, y = imdb_rating)) +
    geom_point() +
    theme_bw() +
    geom_smooth(method = "lm") 

```

The regression line through these points does have a positive slope, however 
the confidence bands clearly contain lines that go anywhere from flat to negative.

This scatterplot is a bit too messy to really discern a trend, so let's look
at deciles of Michael share instead.

```{r}
#Going back to use the full dataset of episodes, including ones Michael isn't in
michael_decile <- michael_ratings %>%
    mutate(decile = cut(michael_share, seq(0, 1, by = .1), right = FALSE))

ggplot(michael_decile, aes(x = decile, y = imdb_rating)) +
    geom_boxplot() +
    theme_bw() +
    labs(x = "Decile of Michael Share")

```

It does seem that the IMDB ratings generally trend upward as Michael speaks more,
though it's not a perfect pattern. Once Michael starts to take over the episode by
speaking up close to 50% or more of total words, the episodes tend to receive lower
ratings.

Does this align with your expectations? Do you wish Michael would
speak **ALL** of the lines in an episode?

Feel free to let me know what you think of this analysis at ryanpeiff@gmail.com.
You can find the full RMarkdown code on my [github](https://github.com/ryanpeiffer/tidytuesday/blob/master/2020_12_The_Office.Rmd).

